---
title: "Practicum II, Part 2: Explore and Mine Data"
author: "Rupert Simpson & Paula Basswerner"
date: "Summer Full 2023"
output: html_notebook
---

### Connect to Database
```{r = datbaseConnection}
# Load required libraries
library(ggplot2)
library(RMySQL)
library(RSQLite)

# Assign remote database values
host <- "sql9.freemysqlhosting.net"
username <- "sql9637087"
password <- "F5WLHRRTXU"
dbname <- "sql9637087"

# Connect to MySQL
mydbcon <- dbConnect(
  MySQL(),
  user = username,
  password = password,
  dbname = dbname,
  host = host
)
```


### Analytical Query I
Top five sales reps with the most sales broken down by year.
```{r = topRepsQueries}
# Query selecting the top 5 representatives per year with the most sales 
top_five_reps_per_year <- "
  SELECT year, firstName, lastName, SUM(total_sold) AS total_sales
  FROM rep_facts
  GROUP BY year, firstName, lastName
  ORDER BY year, total_sales DESC
  LIMIT 5;
"
top_five_reps_per_year.df <- dbGetQuery(mydbcon, top_five_reps_per_year)

# Query selecting the top 5 representatives per quarter for 2020 with the most sales 
top_five_reps_per_quarter <- "
  SELECT quarter, firstName, lastName, SUM(total_sold) AS total_sales
  FROM rep_facts
  GROUP BY quarter, firstName, lastName
  ORDER BY quarter, total_sales DESC;
"
top_five_reps_per_quarter.df <- dbGetQuery(mydbcon, top_five_reps_per_quarter)

```

```{r = dataForReps}
print(top_five_reps_per_year.df)
print(top_five_reps_per_quarter.df)
```

### Visualization for Top Representatives with the most sold per year
This graphs shows the top 5 representatives with the highest sales for the year of 2020.
```{r = topRepsYear}
library(scales)

ggplot(top_five_reps_per_year.df, aes(x = year, y = total_sales, fill = paste(firstName, lastName))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year", y = "Total Sales", title = "Top 5 Sales Reps - Total Sales per Year", fill = "Sales Rep") +
  theme_minimal() +
  scale_y_continuous(labels = comma) 
```

### Visualization for Top Representatives with the most sold per quarter
This graphs shows the top 5 representatives with the highest sales for the year of 2020, arranged by quarter
```{r = topRepsPerQuarter}
ggplot(top_five_reps_per_quarter.df, aes(x = quarter, y = total_sales, color = paste(firstName, lastName))) +
  geom_line() +
  geom_point() +
  labs(x = "Quarter", y = "Total Sales", title = "Top Sales Reps - Total Sales per 2020 Quarter", color = "Sales Rep") +
  theme_minimal()
```

### Analytical Query II
Total revenue sold per month and per year 
```{r = totalSoldPerMonthQueries}

total_sold_per_year <- "
  SELECT year, SUM(total_sold) AS total_sold_per_year
  FROM product_facts
  GROUP BY year;
"

total_sold_per_month <- "
  SELECT month, SUM(total_sold) AS total_sold_per_month
  FROM product_facts
  GROUP BY month;
"

total_sold_per_quarter <- "
  SELECT quarter, SUM(total_sold) AS total_sold_per_quarter
  FROM product_facts
  GROUP BY quarter;
"

total_sold_per_year.df <- dbGetQuery(mydbcon, total_sold_per_year)
total_sold_per_quarter.df <- dbGetQuery(mydbcon, total_sold_per_quarter)
total_sold_per_month.df <- dbGetQuery(mydbcon, total_sold_per_month)
```

### Visualization for Total Sales per Year
This graph shows the total sales per year, in this case year 2020
```{r = graphSalesPerYear}
ggplot(total_sold_per_year.df, aes(x = year, y = total_sold_per_year)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "Total Sold", title = "Total Sold per Year") +
  theme_minimal()
```

### Visualization for Total Sales per Quarter
This graph shows the total sales per quarter 
```{r = graphSalesPerQuarter}
ggplot(total_sold_per_quarter.df, aes(x = quarter, y = total_sold_per_quarter)) +
  geom_line() +  # This adds the line connecting plot points
  geom_point() +
  labs(x = "Quarter", y = "Total Sold", title = "Total Sold per Quarter") +
  theme_minimal()
```

### Visualization for Total Sales per Month 
This graph shows the total sales per month for the avaliable year of 2020 
```{r = graphSalesPerMonth}
ggplot(total_sold_per_month.df, aes(x = month, y = total_sold_per_month)) +
  geom_line() +  # This adds the line connecting plot points
  geom_point() +
  labs(x = "Month", y = "Total Sold", title = "Total Sold per Month") +
  theme_minimal()
```


### Database Disconnect 
```{r = databaseDisconnect}
dbDisconnect(mydbcon)
```